---
title: "Problem Sheet 1"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  Describe each of the following measures of location including their the pros and cons

i Mean

The mean is the point at which the sum of the deviations is 0:

Pros:


* Easy to calculate;

* Uses all the data.

Con:

* Sensitive to extreme values.
 


ii Median

The median is the middle value of the ordered set of values:

Pros:

* Median gives the center of the data;

* Not sensitive to extreme values.

Con:

* Does not use all the data.
	 


iii) Variance

The variance is the spread around the mean:

Pro:

* Takes all data into account.

Cons:


* Hard to interpret;

* Can be influenced by extreme values.
	


iv Skewness


Skewness is a measure of symmetry of a distribution:
Pro

* Takes all data into account;

Con

* Not intuitive.


## Counting

2. How many different combinations of 4 cards can be made for a 52 card deck.

__ANSWER:__

Order __does not__ matter
$$C_4^{52}=\big(\begin{array}{c} 52\\ 4 \end{array} \big)=\frac{52!}{(52-4)!4!}=\frac{(52\times51\times50\times49)}{(4\times3\times2\times1)}=270,725 $$


```{r Q2}
52*51*50*49/(4*3*2*1)

# OR
choose(52,4)

```
Order __does__ matter


__ANSWER:__


$$P_4^{52}=\frac{52!}{(52-4)!}=(52\times51\times50\times49)=6,497,400 $$

```{r Q2ii}
52*51*50*49

# OR
factorial(52)/factorial(48)

```
3. A bank issues bank cards with PINs consisting of 4 digits, each one {0,1,2,…,9}. How many unique PINs are there if
i. Any 4-digit code can be used.

__ANSWER:__


$$10^4=10\times 10 \times 10\times 10\times 10=10,000$$
```{r Q3i}
10*10*10*10
```

ii. The digits must be different.

__ANSWER:__


Order __does__ matter
$$P_4^{10}=\frac{10!}{(10-4)!}=(10\times9\times8\times7)=5,040 $$
```{r Q3ii}
10*9*8*7
```
4.  In a lottery, each ticket has 5 one-digit numbers 0-9 which is not repeated on it.
i You win if your ticket has the digits in any order. What are the total number of possible combinations?

__ANSWER:__


Order __does not__ matter
$$C_5^{10}=\big(\begin{array}{c} 10\\ 5 \end{array} \big)=\frac{10!}{(10-5)!5!}=\frac{(10\times9\times8\times7\times6)}{(5\times4\times3\times2\times1)}=252 $$
```{r Q4i}
choose(10,5)
10*9*8*7*6/(5*4*3*2*1)

```
ii You would win only if your ticket has the digits in the required order. What are the total number of combinations?

__ANSWER:__


Order __does__ matter
$$P_5^{10}=\frac{10!}{5!}=10\times9\times8\times7\times6=30240 $$
```{r Q4ii}
10*9*8*7*6

```
5.  How many different combinations of 6 cards can be made for a 52 card deck if
i) order matters.


__ANSWER:__


Order __does__ matter
$$P_6^{52}=\frac{52!}{(52-6)!}=52\times51\times50\times49\times48\times47=14,658,134,400 $$
```{r Q5i}
52*51*50*49*48*47
factorial(52)/factorial(46)
```
ii) order does not matter

__ANSWER:__


Order __does not__ matter
$$C_6^{52}=\big(\begin{array}{c} 52\\ 6 \end{array} \big)=\frac{52!}{(52-6)!6!}=\frac{(52\times51\times50\times49\times48\times47)}{(6\times5\times4\times3\times2\times1)}=20,358,520 $$
```{r Q5ii}
choose(52,6)
52*51*50*49*48*47/(6*5*4*3*2*1)

```

6. A poker hand consists of 7 cards: 
i)    How many different hands are possible, if order does not matter,

__ANSWER:__


Order __does not__ matter
$$C_7^{52}=\big(\begin{array}{c} 52\\ 7 \end{array} \big)=\frac{52!}{(52-7)!7!}=\frac{(52\times51\times50\times49\times48\times47\times46)}{(7\times6\times5\times4\times3\times2\times1)}=133,784,560 $$
```{r Q6i}
choose(52,7)
52*51*50*49*48*47*46/(7*6*5*4*3*2*1)

```

ii)    How many hands can be made with at least one king and one queen.

__ANSWER:__


$$\frac{4\times4\times(50\times49\times48\times47\times46)}{(5\times4\times3\times2\times1)}=33,900,160 $$
```{r Q6ii}
4*4*50*49*48*47*46/(5*4*3*2*1)

```

7. In a game of 5 card poker what are the number of different possible hands are there?
i) A hand with a pair

__ANSWER:__


$$\frac{52\times3}{2\times1}\times\frac{(48\times47\times46)}{(3\times2\times1)}=1,349,088$$
```{r Q7i}
(52*3)/(2*1)*48*47*46/(3*2*1)

```

aa)A hand with a only pair

__ANSWER:__


$$\frac{52\times3}{2\times1}\times\frac{(48\times44\times40)}{(3\times2\times1)}=1,098,240$$
```{r Q7a}
(52*3)/(2*1)*48*44*40/(3*2*1)

```

 b)A hand with two pair
 

__ANSWER:__


$$\frac{\frac{52\times3}{2\times1}\times\frac{48\times3}{2\times1}}{2\times1}\times\frac{44}{1}=123,552$$
```{r Q7b}
((52*3)/(2*1)*48*3/(2*1))/2*44/1
```
c)A hand with Three of a kind

__ANSWER:__


$$\frac{52\times3\times2}{3\times2\times1}\times\frac{48\times3}{2\times1}=3,744$$
```{r Q7c}
((52*3*2)/(3*2*1)*48*3/(2*1))
```

d)A hand with a Flush (all the same suit)


__ANSWER:__


$$\frac{52\times12\times11\times10\times9}{5\times4\times3\times2\times1}=5,148$$
```{r Q7d}
((52*12*11*10*9)/(5*4*3*2*1))
```

