---
title: "Linear Regression"
output:
  pdf_document: default
---
A linear regression is used to model a linear relationship of the dependent variable $y$ and the regressors $x_1$, $x_2$, ...
	\[  y=\beta_0+\beta_1 x_{1} +\beta_2 x_{2}+..., \]
	where $\beta_0$, $\beta_1$ are the slopes of the regressors.

	
## Height Example
A simple linear regression (correlation) is used to predict the height of 744 children $y$ using the height of their parent $x$,
```{r}
MYPEARSON<-read.csv("PearsonLeeSimple.csv")
Pearson_child_parent<-lm(child~parent,data=MYPEARSON)
summary(Pearson_child_parent)
```


```{r}

library(ggplot2)

ggplot(MYPEARSON, aes(x=parent, y=child)) +ylab("Child Height (in)")+xlab("Parent Height (in)")+
    geom_point(shape=1) +
    scale_colour_hue(l=50) + # Use a slightly darker palette than normal
    geom_smooth(method=lm,   # Add linear regression lines
                se=FALSE)+    # Don't add shaded confidence region
    scale_color_discrete(name = "Parent")

# ggsave("Linear_Regression.png",dpi=300, width = 4, height = 2)
```

## Interpreting the slope of the regression, $\beta$

1. If $\beta$ is close to $0$ then this would suggest that there is little to no relationship between the variable $y$ and the regressor $x$, for extreme example, someones height ($y$) does not correlation with their teachers height $x$.

2. If $\beta$ is greater than $0$ then this means that there is a positive correlation, for example, a tall person $y$ would have a tall parent $x$.

