---
title: "Problem Sheet 2b - Probability Mass Distributions"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggpubr)
library(knitr)
```

## Question 1
1. There are 30 candy covered chocolates in a bag M&M’s. There is a .1 probability that that the candy is red. If X is the number of red M&M’s in the bag.

i Give the binomial probability mass function for X.

__ANSWER:__

\[\Pr(k)= \left( \begin{array}{c}
	30  \\
	k  \\
	\end{array} \right)(0.1)^k(0.9)^{30-k}, \ \ k=0,1,2,...30, \]
	where k is the number of red M & Ms.
	

ii Find the probability of less than 2 red M&Ms in the bag.
 
__ANSWER:__


```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}
n<-30
k<-1
MandM<-0:n # Number of Games
Pr_Red<-0.1 # Probability of Success
Pr_Reds<-c(dbinom(0:n,n,Pr_Red)) # Binomial Probability 
Q1ii<-c(dbinom(0:k,n,Pr_Red))
E_X<- n*Pr_Red # Expected Value
Var_X<-10*Pr_Red*(1-Pr_Red) # Variance

```

\[\Pr(0)= \left( \begin{array}{c}
	30  \\
0  \\
	\end{array} \right)(0.1)^0(0.9)^{30-0}= `r Q1ii[1]`\]
\[\Pr(1)= \left( \begin{array}{c}
	30  \\
1  \\
	\end{array} \right)(0.1)^1(0.9)^{30-1}= `r Q1ii[2]`\]

\[Pr(<2)=Pr(0)+Pr(1)=`r Q1ii[1]`+`r Q1ii[2]`=`r sum(Q1ii)`,  \]

As it is a Binomial Distribution we can state that the expected number of Red is
\[E[k]=`r n`\times `r Pr_Red` =`r E_X`,\]
the variance of the distribution is 
\[Var[k]=`r n`\times `r Pr_Red`\times(1-`r Pr_Red`) =`r Var_X`.\]

The plot below shows the Binomial Distribution M & Ms:

```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}
df <- data.frame(MandM,Pr_Reds)
## Plot
binomial.p_dist<-ggplot(df, aes(x=MandM,y=Pr_Reds)) + geom_col(width = 0.1)+
  xlab("Number of Red M and Ms  out of 30")+
  ylab("Pr of Red")+ggtitle("Binomial Distribution")+
  scale_x_continuous(breaks=0:n)+ 
  geom_vline(xintercept = E_X, color = "Red", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_Reds)),vjust=2, 
            colour="red",  text=element_text(size=18))

 binomial.p_dist
ggsave("Q1.png",dpi=300, width = 4, height = 2)
```


## Question 2

2. A baby wakes on average 0.25 times every hour.
 
i If X is the number of times a baby wakes in an hour, give the poisson probability mass function for X.
__ANSWER:__

The distribution is described by the average, $\lambda=0.25$,
\[\Pr(k)=\frac{0.25^ke^{-0.25}}{k!}, \ \ k=0,1,2,... \]
where $k$ is the number of times the baby wakes every hour.

The expected value of the Poisson Distribution is
\[E[k]=0.25, \]
with the variance
\[Var[k]=0.25. \]



The plot below shows the Poisson Distribution for $\lambda=0.25$ average number of times a baby wakes in an hour:

```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}
Wake<-0:20 # Number of Games
Baby_wakes<-0.25 # Lambda
Pr_wakes<-c(dpois(Wake,Baby_wakes)) # Poisson Probabilities
E_X<-Baby_wakes # Expected Outcome

df <- data.frame(Wake,Pr_wakes)

## Plot
Poisson.p_dist<-ggplot(df, aes(x=Wake,y=Pr_wakes)) + 
  geom_col(width = 0.1)+xlab("Number of Times a Baby wakes in an hour")+
  ylab("Pr of baby Wakes")+ggtitle("Poisson Distribution")+ 
  geom_vline(xintercept = E_X,  color = "blue", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_wakes)),vjust=2,
            colour="blue",text=element_text(size=18))

Poisson.p_dist
ggsave("Q2i.png",dpi=300, width = 4, height = 2)

```






ii If X is the number of times a baby wakes in eight hour, give the poisson probability mass function for X.

__ANSWER:__

The distribution is described by the average, $\lambda=0.25\times 8 =2$,
\[\Pr(k)=\frac{2^ke^{-2}}{k!}, \ \ k=0,1,2,... \]
where $k$ is the number of times the baby wakes every 8 hours.

The expected value of the Poisson Distribution is
\[E[k]=2, \]
with the variance
\[Var[k]=2. \]


The plot below shows the Poisson Distribution for $\lambda=2$ average number of times a baby wakes in eight hours:

```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}

Baby_wakes<-0.25*8 # Lambda
Pr_wakes<-c(dpois(Wake,Baby_wakes)) # Poisson Probabilities
E_X<-Baby_wakes # Expected Outcome

df <- data.frame(Wake,Pr_wakes)

## Plot
Poisson.p_dist<-ggplot(df, aes(x=Wake,y=Pr_wakes)) + 
  geom_col(width = 0.1)+xlab("Number of Times a Baby wakes in an hour")+
  ylab("Pr of baby Wakes")+ggtitle("Poisson Distribution")+ 
  geom_vline(xintercept = E_X,  color = "blue", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_wakes)),vjust=2,
            colour="blue",text=element_text(size=18))

Poisson.p_dist
ggsave("Q2i.png",dpi=300, width = 4, height = 2)

```






iii What is the probability that the baby does not wake during the 8 hours.

__ANSWER:__

```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}

Baby_wakes<-0.25*8 # Lambda
Pr_wakes_zero<-dpois(0,Baby_wakes) #
```

\[\Pr(0)=\frac{0.25^0e^{-0.25}}{0!}=`r Pr_wakes_zero`.  \]

## Question 3

3. Give the features of a
 
i Geometric Experiment. 

__ANSWER:__

• Bernoulli Trial
• Goes until you "win"
• E[X]=1/p, VAR[X]=q/p^2


ii Binomial Experiment. 

__ANSWER:__

• Bernoulli Trial
• Play a specific number of times
• E[x]=np, VAR[X]=npq



iii Poisson Experiment.

__ANSWER:__

• Well known mean
• Number of "wins"
• E[X]=lambda, VAR[X]=lambda


## Question 4

4. Every day a production line makes 100 computers of which 10% are defective. If X is the number of defective computers in a day.

a) Give the binomial probability mass function for X.

__ANSWER:__

\[\Pr(k)= \left( \begin{array}{c}
	100  \\
	k  \\
	\end{array} \right)(0.1)^k(0.9)^{30-k}, \ \ k=0,1,2,...100, \]
	where k is the number of defective computers.
	

b) Find the probability that there is more than 2 computers defective in a day.


__ANSWER:__


```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}
n<-100
k<-2
Computers<-0:n # Number of Games
Pr_Def<-0.1 # Probability of Success
Pr_Defs<-c(dbinom(0:n,n,Pr_Def)) # Binomial Probability 
Q4ii<-c(dbinom(0:k,n,Pr_Def))
E_X<- n*Pr_Def # Expected Value
Var_X<-10*Pr_Def*(1-Pr_Def) # Variance

```

\[\Pr(0)= \left( \begin{array}{c}
	100  \\
0  \\
	\end{array} \right)(0.1)^0(0.9)^{100-0}= `r Q4ii[1]`,\]
\[\Pr(1)= \left( \begin{array}{c}
	100  \\
1  \\
	\end{array} \right)(0.1)^1(0.9)^{100-1}= `r Q4ii[2]`,\]

\[Pr(2)= \left( \begin{array}{c}
	100  \\
2  \\
	\end{array} \right)(0.1)^2(0.9)^{100-2}= `r Q4ii[3]` .\]

\[ Pr(>2)=1-Pr(\leq2)=1-(Pr(0)+Pr(1)+Pr(2))=1-`r Q4ii[1]`+`r Q4ii[2]`+`r Q4ii[3] ` = `r 1-sum(Q4ii)`  \]

The plot below shows the Binomial Distribution for defective computers:

```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}
df <- data.frame(Computers,Pr_Defs)
## Plot
binomial.p_dist<-ggplot(df, aes(x=Computers,y=Pr_Defs)) + geom_col(width = 0.1)+
  xlab("Number of Defective Computers")+
  ylab("Pr of Defective")+ggtitle("Binomial Distribution")+
  geom_vline(xintercept = E_X, color = "Red", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_Defs)),vjust=2, 
            colour="red",  text=element_text(size=18))

 binomial.p_dist
ggsave("Q4.png",dpi=300, width = 4, height = 2)
```

## Question 5
5. A phone center receives 15 calls every 30 minutes.
 
a) If X is the number of phone calls in 30 minutes, give the Poisson probability mass function for X.

__ANSWER:__

The distribution is described by the average, $\lambda=15$,
\[\Pr(k)=\frac{15^ke^{-15}}{k!}, \ \ k=0,1,2,... \]
where $k$ is the number of phone calls per half hour.

The expected value of the Poisson Distribution is
\[E[k]=15, \]
with the variance
\[Var[k]=15. \]



The plot below shows the Poisson Distribution for $\lambda=15$ average number of calls per half-hour:

```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}
Calls<-0:50 # Number of Games
No_of_Calls<-15 # Lambda
Pr_Calls<-c(dpois(Calls,No_of_Calls)) # Poisson Probabilities
E_X<-No_of_Calls # Expected Outcome

df <- data.frame(Calls,Pr_Calls)

## Plot
Poisson.p_dist<-ggplot(df, aes(x=Calls,y=Pr_Calls)) + 
  geom_col(width = 0.1)+xlab("Number of Phone Calls in a half hour")+
  ylab("Pr of Phone Calls")+ggtitle("Poisson Distribution")+ 
  geom_vline(xintercept = E_X,  color = "blue", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_Calls)),vjust=2,
            colour="blue",text=element_text(size=18))

Poisson.p_dist
ggsave("Q5i.png",dpi=300, width = 4, height = 2)

```



b) What is the probability that there will be exactly 10 phone calls in the first 30 minutes and exactly 20 phone calls in the second 30 minutes.

__ANSWER:__


```{r,echo=TRUE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}
No_of_Calls<-15 # Lambda
Pr_Calls_10<-dpois(10,No_of_Calls) #  10 Calls Poisson Probability
Pr_Calls_20<-dpois(20,No_of_Calls) #  20 Calls Poisson Probability
Pr_Calls_10_then_20 <- Pr_Calls_10*Pr_Calls_20
```

Probability of 10 calls in the first half-hour:
\[\Pr(10)=\frac{15^10e^{-15}}{10!}=`r Pr_Calls_10`. \]
Probability of 20 calls in the first half-hour:

\[\Pr(20)=\frac{15^20e^{-15}}{20!}=`r Pr_Calls_20`. \]

The combination of both:
\[\Pr(10)\times \Pr(20)=`r Pr_Calls_10_then_20`. \]


