---
title: 'Discrete Distributions'
author: '[John S Butler (TU Dublin)](https://johnsbutler.netlify.com)'
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggpubr)

```
# Introduction



# Binomial Distribution
A Binomial Distribution is used to describe the probability distribution of $n$ experiments that have two possible outcomes "success" or "failure".
The probability of "success" is $p$, the probability of failure is $q=1-p$.
This gives the general deifintion of the distribution as:
\[\Pr(k)= \left( \begin{array}{c}
	n  \\
	k  \\
	\end{array} \right)p^kq^{n-k}, \ \ k=0,1,2,...n, \]
where $k$ is the number of "successes". 
If the data can be described using a Binomial Distribution then the expected outcome, $E[k]$ of the distribution is given by:
\[E[k]=np,\]
and the variance of the distribution $Var[k]$, is 
\[Var[k]=npq.\]
	
### Example
The example we shall use to illustrate the discrete distibutions is the New Zealand vs Ireland Wolrd Cup Rugby Quarter Final.
Let's assume the probability of "success", Ireland beating New Zealand is 
\[p=0.15,\]
the probability of failure is 
\[q=1-p=1-0.15=0.85.\]
Now imagine that Ireland have to play New Zealand 10 times, this gives the Binomial Distribution
\[\Pr(k)= \left( \begin{array}{c}
	10  \\
	k  \\
	\end{array} \right)(0.15)^k(0.85)^{10-k}, \ \ k=0,1,2,...10, \]
	where k is the number of times Ireland beats New Zealand.
As it is a binomial Distribution we can state that the expected number of times Ireland will win is
\[E[k]=10\times0.15=1.5,\]
ie Ireland is only expected to win once, the variance of the distribution is 
\[Var[k]=10\times 0.15 \times 0.85=1.35.\]



```{r,echo=FALSE}
Games<-0:10
Pr_ireland_win<-0.15
Pr_win<-c(dbinom(0:10,10,Pr_ireland_win))
Cum_Pr_win<-c(pbinom(0:10,10,Pr_ireland_win))
E_X<- 10*Pr_ireland_win
Var_X<-10*Pr_ireland_win*(1-Pr_ireland_win)

df <- data.frame(Games,Pr_win,Cum_Pr_win)
binomial.p_dist<-ggplot(df, aes(x=Games,y=Pr_win)) + geom_col(width = 0.1)+xlab("Number of Irish Wins \n out of 10 games")+ylab("Pr of Wins")+ggtitle("Binomial \n Distribution")+scale_x_continuous(breaks=0:10)+
theme(text = element_text(size=8))+ geom_vline(xintercept = E_X, 
                color = "blue", size=1)+geom_vline(xintercept = E_X-Var_X, line="dashed",color = "blue", size=1)+
  geom_vline(xintercept = E_X+Var_X, line="dashed",color = "blue", size=1)+geom_text(aes(x=E_X+0.1, label="E[k]", y=max(Pr_win)),vjust=2, colour="blue", angle=0, vjust = 1.2, text=element_text(size=18))

binomial.cum_p_dist<-ggplot(df, aes(x=Games,y=Cum_Pr_win)) + geom_col(width = 0.1)+ylab("Pr(<= Irish Wins)")+xlab("Number of Irish Wins \n out of 10 games")+ggtitle("Cummulative Binomial \n Distribution")+scale_x_continuous(breaks=0:10)+
theme(text = element_text(size=8))

 binomial.p_dist
#ggsave("Binomial.png",dpi=300, width = 4, height = 2)

```

# Geometric Distribution

## Definition:

\[\Pr(k)=q^{(k-1)}p, \ \ k=1,2,... \]

\[E[k]=\frac{1}{p}, \ \ \ Var[k]=\frac{q}{p^2}.\]
	
In R code:



```{r,echo=FALSE}
Games<-0:20
Pr_ireland_win<-0.1
Pr_win<-c(dgeom(Games,Pr_ireland_win))
Cum_Pr_win<-c(pgeom(Games,Pr_ireland_win))

df <- data.frame(Games,Pr_win,Cum_Pr_win)
binomial.p_dist<-ggplot(df, aes(x=Games,y=Pr_win)) + geom_col(width = 0.1)+xlab("Number of Games until\n Ireland Win")+ylab("Pr(Games until Ireland Win)")+ggtitle("Geometric \n Distribution")+scale_x_continuous(breaks=Games)+
theme(text = element_text(size=8))#+scale_x_discrete(breaks=0:10)

binomial.cum_p_dist<-ggplot(df, aes(x=Games,y=Cum_Pr_win)) + geom_col(width = 0.1)+xlab("Number of Games until\n Ireland Win")+ylab("Pr <= Wins")+ggtitle("Cummulative Geometric\n Distribution")+scale_x_continuous(breaks=Games)+
theme(text = element_text(size=8))
ggarrange( binomial.p_dist,binomial.cum_p_dist, ncol = 2, nrow = 1)
#ggsave("Geometric.png",dpi=300, width = 4, height = 2)

```


# Poisson Distribution

## Definition:

\[\Pr(k)=\frac{\lambda^ke^{-\lambda}}{k!}, \ \ k=0,1,2,... \]

\[E[k]=\lambda, \ \ \ Var[k]=\lambda. \]
	
In R code:



```{r,echo=FALSE}
Games<-0:10
Pr_ireland_win<-1
Pr_win<-c(dpois(0:10,Pr_ireland_win))
Cum_Pr_win<-c(ppois(0:10,Pr_ireland_win))

df <- data.frame(Games,Pr_win,Cum_Pr_win)
binomial.p_dist<-ggplot(df, aes(x=Games,y=Pr_win)) + geom_col(width = 0.1)+xlab("Number of Irish Wins in a \n 15 year period")+ylab("Pr of Irish Wins")+ggtitle("Poisson \n Distribution")+scale_x_continuous(breaks=0:10)+
theme(text = element_text(size=8))#+scale_x_discrete(breaks=0:10)
binomial.cum_p_dist<-ggplot(df, aes(x=Games,y=Cum_Pr_win)) + geom_col(width = 0.1)+xlab("Number of Irish Wins in a \n 15 year period")+ylab("Pr <= number of Irish Wins")+ggtitle("Cummulative Poisson \n Distribution")+scale_x_continuous(breaks=0:10) +
theme(text = element_text(size=8))
ggarrange( binomial.p_dist,binomial.cum_p_dist, ncol = 2, nrow = 1)
#ggsave("Poisson.png",dpi=300, width = 4, height = 2)

```

# Normal Distribution

## Definition:

\[\Pr(k)=\frac{\lambda^ke^{-\lambda}}{k!}, \ \ k=0,1,2,... \]

\[E[k]=\lambda, \ \ \ Var[k]=\lambda. \]
	
In R code:



```{r,echo=FALSE}
Z<-seq(-4,4,by=0.1)
Pr_ireland_win<-0.1
Pr_win<-c(dnorm(Z))
Cum_Pr_win<-c(pnorm(Z))

df <- data.frame(Z,Pr_win,Cum_Pr_win)
binomial.p_dist<-ggplot(df, aes(x=Z,y=Pr_win)) + geom_line(width = 0.1)+xlab("Z-score")+ylab("Pr of Z ")+ggtitle("Gaussian (Normal) \n Distribution")+
theme(text = element_text(size=8))#+scale_x_discrete(breaks=0:10)

binomial.cum_p_dist<-ggplot(df, aes(x=Z,y=Cum_Pr_win)) + geom_line(width = 0.1)+ylab("Pr(<= Z)")+xlab("Z -score")+ggtitle("Cummulative Gaussian \n Distribution")+
theme(text = element_text(size=8))

ggarrange( binomial.p_dist,binomial.cum_p_dist, ncol = 2, nrow = 1)
#ggsave("Normal.png",dpi=300, width = 4, height = 2)
```


## Confidence Intervals


```{r,echo=FALSE}

ggplot(df, aes(x=Z,y=Pr_win)) + geom_line(size=1)+ylab("Pr(Z)")+xlab("Z-score")+
                                ggtitle("Confidence Intervals")+
                                theme(text = element_text(size=8))+
                                geom_vline(xintercept = 1.96, linetype="dotted",color = "darkgray", size=0.5)+
                                geom_vline(xintercept = -1.96, linetype="dotted",color = "darkgray", size=0.5)+
                                geom_vline(xintercept = 2.58, linetype="dashed", color = "black", size=0.5)+
                                geom_vline(xintercept = -2.58, linetype="dashed",color = "black", size=0.5)+
                                geom_segment(aes(x = -1.96, y = 0.42, xend = 1.96, yend = 0.42),linetype="dotted",color="darkgray")+
                                geom_segment(aes(x = -2.58, y = 0.5, xend = 2.58, yend = 0.5),linetype="dashed")+
                                geom_text(aes(x=0, label="95% Confidence Interval",y=0.46),vjust=1,size=2,color="darkgray")+
                                geom_text(aes(x=0, label="99% Confidence Interval",y=0.54),vjust=1,size=2)


#ggsave("CI_Normal.png",dpi=300, width = 4, height = 2)
```
