---
title: "Chi-squared Hypothesis Testing"
output: word_document
---
# Chi-squared Tests
## Goodness of Fit

### Example 1 
Use the Chi-squared test to investigate if the observed data is different from the predicted data

+ i For a normal distribution with mean 170cm and mean 10cm

```{r,echo=FALSE, warning=FALSE}
library(knitr)
library(ggplot2)
Bins<-c("<160" ,"160-170" ,"170-180", "180-")
Observed<-c(3,3,7,2)
Expected<-c(2.4,5.1,5.1,2.4)

df<-data.frame(Bins,Observed,Expected)
kable(df,label = "test")

ggplot(df,aes(x=Bins))+geom_point(aes(y=Observed),shape=15,color="red",size=3)+geom_point(aes(y=Expected),shape=16,size=3,fill="blue")+geom_line(aes(x=Bins,y=Expected),color="blue")

chisq.test(Observed, p = Expected/sum(Expected))

```

+ ii For a flat distribution

```{r,echo=FALSE, warning=FALSE}
Expected<-c(3.75, 3.75, 3.75, 3.75)
df<-data.frame(Bins,Observed,Expected)
kable(df)
chisq.test(Observed, p = Expected/sum(Expected))

```
## Test of Independence 
An ice-cream company had 500 people sample one of three different ice-cream 
flavours and asked them to say whether they liked or disliked the ice-cream. The resulting observed data is presented in the table below



```{r,echo=FALSE,warning=FALSE}

Icecream<-matrix(c(130,170,100,20,30,50),ncol =3,byrow=TRUE)
rownames(Icecream)<-c("Liked","Disliked")
colnames(Icecream)<-c("Vanilla","Chocolate","Strawberry")
Icecream
```
Conduct a hypothesis test to determine if these data supply evidence (alpha = 0.05) that the enjoyment of the ice-cream depends on the flavour

```{r,echo=FALSE,warning=FALSE}

chisq.test(Icecream)

```