---
title: "Hypothesis Testing Solutions"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggpubr)
library(knitr)
```

# Z- test

## Example 1
The mean height from the entire of men and women in the population is 170cm. You
want to check if there is a statistically significant difference (with a significance level of
95%) between the means of the sample and the population.
Calculate the z-test to test is the observed data is different from the population data for

+ i 10 men of height: 179, 188, 172, 185, 171, 199, 176, 188, 158, 200

```{r,dpi=300,echo=FALSE, warning=FALSE}
data<-c(179, 188, 172, 185, 171, 199, 176, 188, 158, 200)

t.test(data,mu=170)

df <- data.frame(data)

ggboxplot(df,  y = "data",
          add = "jitter")+geom_hline(yintercept = 170)

```

+ ii 12 women of height: 175, 158, 159, 167, 171, 151, 160, 167, 147, 172, 154, 152
```{r,dpi=300,echo=FALSE, warning=FALSE}
data<-c(175, 158, 159, 167, 171, 151, 160, 167, 147, 172, 154, 152)

t.test(data,mu=170)

df <- data.frame(data)

ggboxplot(df,  y = "data",
          add = "jitter")+geom_hline(yintercept = 170)


```
+ iii 7 women and 8 men of height: 166, 189, 180, 193, 148, 174, 161, 170, 172, 179, 158, 157, 176, 171, 174
```{r,dpi=300,echo=FALSE, warning=FALSE}
data<-c(166, 189, 180, 193, 148, 174, 161, 170, 172, 179, 158, 157, 176,171, 174)

t.test(data,mu=170)

df <- data.frame(data)

ggboxplot(df,  y = "data",
          add = "jitter")+geom_hline(yintercept = 170)


```

+ iv Compare the height of the 10 men and the 12 women

```{r,dpi=300,echo=FALSE, warning=FALSE}
men<-c(179, 188, 172, 185, 171, 199, 176, 188, 158, 200)
women<-c(175, 158, 159, 167, 171, 151, 160, 167, 147, 172, 154, 152)
t.test(men,women,paired=FALSE)

Sex<-c(rep("M",length(men)),rep("W",length(women)))
Height<-c(men,women)
df <- data.frame(Sex,Height)

ggboxplot(df, x = "Sex", y = "Height",
          color = "Sex", palette = "jco",
          add = "jitter")

```


### QUESTION 2. 
A survey claims that 8 out of 10 doctors recommend aspirin for their patients with headaches. To test this claim, a random sample of 100 doctors is obtained. Of these 95 doctors, 82 indicate that they recommend aspirin.

Calculate the z-test to test is the observed data different from the population data.

```{r,dpi=300,echo=FALSE, warning=FALSE}
p0<-0.8

q0<-1-p0
p<-82/95
n<-95
Z<-(p-p0)/sqrt(p0*q0/n)
Z
```

# t-test
### QUESTION 3 
Read in the PearsonLeeSimple.csv data. Using an unpaired t-test compare:

* the heights of children and parents,
* within in the parents compared the Height of Mothers and Fathers.

```{r,dpi=300,echo=FALSE, warning=FALSE}
JB_PEARSON<-read.csv("PearsonLeeSimple.csv")

 t.test(JB_PEARSON$parent,JB_PEARSON$child,paired=FALSE)
 t.test(JB_PEARSON$parent~JB_PEARSON$par,paired=FALSE)



ggboxplot(JB_PEARSON, x="par", y = "parent",
          add = "jitter",color="par")

```


