---
title: 'Measures of Location: First Moment'
author: '[John S Butler (TU Dublin)](https://johnsbutler.netlify.com)'
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

Different aspects of a distribution of data can be summarised by the measures of location:

1. The First Moment: Middle.

2. The Second Moment: Spread.

3. The Third Moment: Symmetry.


All that being said, I would always recommend plotting the data first before anything else.

__A picture (histogram) is worth a thousand words.__


# First Moment: Middle

## Mean (average)
### Definition:

The mean $\bar{x}$, $\mu$ is the point around which the sum of the deviations is 0.
The formula is given by
$$ \bar{x}=\frac{\sum_{i=1}^n x_i}{n} $$
where $x_i$ is each element and $n$ is the number of elements.
In words, add the numbers together and divided by how many numbers are in the list.

### Example
Given the list of 7 ages at a concert {19,18,20,18,18,18,20}, the mean is:

$$ \bar{x}=\frac{20+18+19+18+18+18+20}{7}=18.71429. $$


In R code:

```{r}
Age=c(20, 18,19, 18,18,18,20) # List of 7 numbers
## Cacluate the mean in the long form
(20+18+19+18+18+18+19+20)/7
## Cacluate the mean using the function mean
mean(Age) # The mean of the list

```
### Graphical representation of the mean:

```{r,warning=FALSE,echo=FALSE}

library(ggplot2)

df <- data.frame(Age)
ggplot(df, aes(x=Age)) +
  geom_histogram(color="black", fill="white",binwidth = 1)+xlab("Age (yrs)")+
  ylab(NULL)+ theme(plot.title = element_text(hjust=1.0))+
geom_vline(xintercept=mean(Age),col="blue",size=1)+
  geom_text(aes(x=mean(Age), label="mean", y=6), colour="blue",  text=element_text(size=11))
ggsave("Mean.png",dpi=300, width = 2, height = 1.5)
```



### Mean Pros and Cons
Pros of the mean:

* The mean is easy to calculate.

* The mean is well known and understood.

Cons of the mean:
* The mean is  senstitive to extreme values.

Given the list of the ages of  8 people at a concert {20,18,19, 18,56, 18,18,20}, the mean is:

$$ \bar{x}=\frac{20+18+19+18+18+56+18+20}{8}=23.375. $$


In R code:

```{r}
Age=c(20,18,19, 18,56, 18,18,20) # List of 8 ages
## Cacluate the mean using the function mean
mean(Age) # The mean of the list
df <- data.frame(Age)
```
## Median

### Definition:

The median is the middle values of an ordered list.

### Example
Given the list of 7 ages at a concert {19,18,20,18,18,18,20}. 

1. First the list has to be ordered
  {18,18,18,18,19,20,20}.
  
2. Then the middle of the list found
  {18,18,18,*18*,19,20,20}.
  
The median is 18.


In R code:

```{r}
Age=c(20,18,19,18,18,18,20) # List of 7 numbers

median(Age)

```

### Graphical representation of the median:

```{r,warning=FALSE,echo=FALSE}

df <- data.frame(Age)
ggplot(df, aes(x=Age)) +  geom_histogram(color="black", fill="white",binwidth = 1)+xlab("Age (yrs)")+ylab(NULL)+ theme(plot.title = element_text(hjust=1.0))+
geom_vline(xintercept=median(Age),col="green",size=1)+
  geom_text(aes(x=median(Age), label="median", y=6), colour="green",  text=element_text(size=11))
ggsave("Median.png",dpi=300, width = 2, height = 1.5)




```

### Median Pros and Cons
Pros of the median:

* The median is the centre of the data - very intuitive.
* Not as sensitive as the mean to extreme values.
 


Cons of median:
* The calculation of the median does not use all data.
* Sorting a long list can be computationally expensive.

## Mode

### Definition:

The mode is the most commonly occurring value in the distribution.

### Example
Given the list of 7 ages at a concert {19,18,20,18,18,18,20}. 
The most frequent values is 18, so 18 in the mode.


### In R code:

```{r}
Age=c(20,18,19,18,18,18,20) # List of 7 numbers
# Function to calculate mode
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
Mode(Age)

```
### Graphical representation of the mode:

```{r,warning=FALSE,echo=FALSE}

df <- data.frame(Age)
ggplot(df, aes(x=Age)) +  geom_histogram(color="black", fill="white",binwidth = 1)+xlab("Age (yrs)")+ylab(NULL)+ theme(plot.title = element_text(hjust=1.0))+
geom_vline(xintercept=Mode(Age),col="red",size=1)+
  geom_text(aes(x=Mode(Age), label="Mode", y=6), colour="red",  text=element_text(size=11))
ggsave("Mode.png",dpi=300, width = 2, height = 1.5)

```


### Median Pros and Cons
Pros of the mode:

* The mode shows where the data is concentrated.

Cons of the mode:

* More than one Mode in a dataset.
* The calculation of the mode does not use all the data.

## All in one plot

Plotting a histogram of Heights from the Pearson dataset

```{r,warning=FALSE,echo=FALSE}

df<-read.csv("PearsonLeeSimple.csv")
attach(df)
ggplot(df, aes(x=child)) +  geom_histogram(color="black", fill="white",binwidth = 1)+xlab("Height (inches)")+ylab(NULL)+ theme(plot.title = element_text(hjust=1.0))+geom_vline(xintercept=mean(child),col="blue",size=1,labels="mean")+
  geom_text(aes(x=mean(child), label="mean", y=90), colour="blue",  text=element_text(size=11))+
geom_vline(xintercept=median(child),col="green",size=1)+
  geom_text(aes(x=median(child), label="median", y=100), colour="green",  text=element_text(size=11))+
geom_vline(xintercept=Mode(child),col="red",size=1)+
  geom_text(aes(x=Mode(child), label="mode", y=80), colour="red",  text=element_text(size=11))
ggsave("Pearson.png",dpi=300, width = 4, height = 2.5)
```

The figure below shows a histogram of 3000 attendees age observations at a Wiggles concert:
```{r,warning=FALSE,echo=FALSE}
Age<-c(rnorm(2000,10,1.5),rnorm(1000,40,6))

df <- data.frame(Age)
ggplot(df, aes(x=Age)) +  geom_histogram(color="black", fill="white",binwidth = 1)+xlab("Age (yrs)")+ylab(NULL)+ theme(plot.title = element_text(hjust=1.0))+geom_vline(xintercept=mean(Age),col="blue",size=1,labels="mean")+
  geom_text(aes(x=mean(Age), label="mean", y=600), colour="blue",  text=element_text(size=11))+
geom_vline(xintercept=median(Age),col="green",size=1)+
  geom_text(aes(x=median(Age), label="median", y=600), colour="green",  text=element_text(size=11))+
geom_vline(xintercept=Mode(Age),col="red",size=1)+
  geom_text(aes(x=Mode(Age), label="mode", y=600), colour="red",  text=element_text(size=11))
ggsave("middle.png",dpi=300, width = 4, height = 2.5)


```

