---
title: 'Measures of Location: Third Moment'
author: '[John S Butler (TU Dublin)](https://johnsbutler.netlify.com)'
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

A different aspects of a distribution of data can be summarised by the measures of location:

1. The First Moment: Middle.

2. The Second Moment: Spread.

3. __The Third Moment: Symmetry.__


All that being said, I would always recommend plotting the data first before anything else.

__A picture (histogram) is worth a thousand words.__



# Third Moment: Symmetry

## Skewness
### Definition 1:
Skewness is a measure of symmetry (or not symmetry) of a distribution.
Pearson's Coefficient of Skewness number 1 uses the mode to calculate skewness, given by the formula is:
$$ sk_1=\frac{\bar{x}-Mode(x)}{\sigma}.$$

### Definition 2:
Pearson's Coefficient of Skewness number 2 uses the median to calculate skewness, given by the formula is:
$$ sk_2=3\frac{\bar{x}-Median(x)}{\sigma}$$,
where $\bar{x}$ is the average of the elements, $\sigma$ is the standard #deviation, $Mode(x)$ is the mode of the elements and $Median(x)$ is the median of the elements.

# Interpretation of Skewness
How to interpret Skewness:

* A skewness near means the distribution is symmetric.

* A Negative skewness means the distribution is right skewed

* A Positive skewness means the distribution is left skewed

# Symmetric Example (Age at a Good Concert)

The figure below shows a symmetric histogram of 2000 concert attendees ages observations at a "good" concert:


```{r,warning=FALSE,echo=FALSE}
Age<-rnorm(2000,23,2)
sk_title = paste0("Symmetric Distribution Sk_1 = ",round((mean(Age)-median(Age))/sd(Age), digits = 3))
#png("Third_Moment_sym.png", width = 1600, height = 800, units = "px", pointsize = 20)

hist(Age,col="lightblue",xlab="Age (years)",xlim=c(15,30),main=sk_title,breaks=100) # Histogram of the 7 numbers
abline(v=mean(Age),col="blue",lwd=3)# line indicating the mean
abline(v=median(Age),col="green",lwd=3)# line indicating the mean
legend("topright",c("mean","median"),lwd=c(3,3),lty=c(1,1),col=c("blue","green"))
#dev.off()
```

The age distribution is equally distributed around the mean, hence the skewness is more or less 0.

# Positive Skewness Example (Age at a Bieber Concert)

The figure below shows a positive skewness histogram of 2000 attendees age observations at a Justin Bieber concert:

```{r,warning=FALSE,echo=FALSE}
Age<-rchisq(2000,7)+15
sk_title = paste0("Postive Skewed Distribution Sk_1 = ", round((mean(Age)-median(Age))/sd(Age), digits = 3))
#png("Third_Moment_pos.png", width = 1600, height = 800, units = "px", pointsize = 20)
hist(Age,col="lightblue",xlab="Age (years)",main=sk_title,breaks=100) # Histogram of the 7 numbers
abline(v=mean(Age),col="blue",lwd=3)# line indicating the mean
abline(v=median(Age),col="green",lwd=3)# line indicating the mean
legend("topright",c("mean","median"),lwd=c(3,3),lty=c(1,1),col=c("blue","green"))
#dev.off()
```

The age distribution shows that the concert attendees are mostly young, hence the skewness is positive.



## Negative Skewed Example (Age at an Andre Rieu Concert)

The figure below shows a negative skewness histogram of 2000 attendees age observations at a Andre Rieu concert:


```{r,warning=FALSE,echo=FALSE}
Age<-rbeta(2000,10,3)*20+70
#png("Third_Moment_neg.png", width = 1600, height = 800, units = "px", pointsize = 20)

sk_title = paste0("Negative Skewed Distribution Sk_1 = ", round((mean(Age)-median(Age))/sd(Age), digits = 3))
hist(Age,col="lightblue",xlab="Age (years)",main=sk_title,breaks=100) # Histogram of the 7 numbers
abline(v=mean(Age),col="blue",lwd=3)# line indicating the mean
abline(v=median(Age),col="green",lwd=3)# line indicating the mean
legend("topleft",c("mean","median"),lwd=c(3,3),lty=c(1,1),col=c("blue","green"))
#dev.off()
```
The age distribution is skewed to the right side as the concert has mostly older adults, hence the skewness is negative.

### All in one plot

```{r,echo=FALSE,warning=FALSE}
library(ggplot2)
library(ggpubr)
Sym<-rnorm(2000,23,2)
Neg<-rbeta(2000,10,3)*20+70
Pos<-rchisq(2000,7)+15
df <- data.frame(Neg,Sym,Pos)
p.neg<-ggplot(df, aes(x=Neg)) +  geom_histogram(color="black", fill="white")+ggtitle("Negative ")+xlab("Age (yrs)")+ylab(NULL)+ theme(plot.title = element_text(hjust=1.0))
p.sym<-ggplot(df, aes(x=Sym)) +  geom_histogram(color="black", fill="white")+ggtitle("Symmetric")+xlab("Age (yrs)")+ylab(NULL)
p.pos<-ggplot(df, aes(x=Pos)) +  geom_histogram(color="black", fill="white")+ggtitle("Positive")+xlab("Age (yrs)")+ylab(NULL)
ggarrange( p.pos,p.sym,  p.neg, ncol = 3, nrow = 1)
ggsave("skew.png",dpi=300, width = 8, height = 6)


```



## Bi-modal distribution
The figure below shows a histogram of 3000 attendees age observations at a Wiggles concert:
```{r,warning=FALSE,echo=FALSE}
Age<-c(rnorm(2000,10,1.5),rnorm(1000,40,6))
sk_title = paste0("Symmetric Distribution Sk_1 = ", round((mean(Age)-median(Age))/sd(Age), digits = 3))
hist(Age,col="lightblue",xlab="Age (years)",main=sk_title,breaks=300) # Histogram of the 7 numbers
abline(v=mean(Age),col="blue",lwd=3)# line indicating the mean
abline(v=median(Age),col="green",lwd=3)# line indicating the mean
legend("topleft",c("mean","median"),lwd=c(3,3),lty=c(1,1),col=c("blue","green"))

```

The distribution is bimodal as there are 2000 excited children with 1000 parents wear earplugs, the skewness is positive but this is misleading.



