---
title: "Conditional Probability"
author: '[John S Butler (TU Dublin)](https://johnsbutler.netlify.com)'
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



The Conditional Probability $\Pr(A|B)$ denotes the probability of the event $A$ occurring given that the event $B$ has occurred,
$$ \Pr(A|B)=\frac{\Pr(A\bigcap B)}{\Pr(B)}.$$

# Example: The rain in Ireland

A normal probability would be what is the probability it is going to rain. $\Pr(\text{rain})$.

A conditional probability would, be what is the probability it is going to rain \textbf{given} that you are in Ireland, $\Pr(\text{rain}|\text{Ireland}),$

$$ \Pr(\text{rain}|\text{Ireland})=\frac{\Pr(\text{rain} \bigcap \text{Ireland})}{\Pr(\text{Ireland})},$$

where the probability of rain is $\Pr(\text{rain})=0.3$, the probability of being in Ireland is $\Pr(\text{Ireland})=0.4)$ and the probability of being in Ireland and it raining is $\Pr(\text{rain}\bigcap \text{Ireland})=0.2$. Given these numbers we can calculate

$$ \Pr(\text{rain}|\text{Ireland})=\frac{0.2}{0.4}=0.5,$$

You could be interested in the probability that you are in Ireland \textbf{given} that it is raining, 
$$ \Pr(\text{Ireland}|\text{rain})=\frac{\Pr(\text{rain}\bigcap \text{Ireland})}{\Pr(\text{rain})}=\frac{0.2}{0.3}.$$
```{r}
Pr_Rain<-0.3
Pr_Ireland<-0.4
Pr_Rain_and_Ireland<-0.2

Pr_Rain_Given_Ireland<-Pr_Rain_and_Ireland/Pr_Ireland
Pr_Rain_Given_Ireland
Pr_Ireland_Given_Rain<-Pr_Rain_and_Ireland/Pr_Rain
Pr_Ireland_Given_Rain

```