---
title: "Mathematical Probability"
author: '[John S Butler (TU Dublin)](https://johnsbutler.netlify.com)'
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Definitions

Define some event $A$ that can be the outcome of an experiment.
$\text{Pr}(A)$ is the probability of a given event $A$ will happen.
Rules:

1. $\text{Pr}(A)$ is between $0$ and $1$, $0\leq Pr(A) \leq 1$;

2. $\text{Pr}(A)=1$, means it will definitely happen;

3. $\text{Pr}(A)=0$, means it will definitely \textbf{not} happen;

4. $\text{Pr}(A)=0.05$, is arbitrarily considered unlikely.
  

# Sample Space and Events
The \textbf{Sample Space}, $S$, of an experiment is the universal set of all possible outcomes for that experiment, defined so, no two outcomes can occur simultaneously. For example:



## Axioms of Probabilities
An event, $A$, is a subset, $\subset$, of the Sample Space
 associated a number $Pr(A)$, the probability of $A$, which must have the following properties

1. If $Pr(A \bigcap B)=0$, then $\text{Pr}(A \bigcup B)= \text{Pr}(A)+\text{Pr}(B)$;

2.  Probability of the Null Event $\text{Pr}(\emptyset)=0$;

3. The probability of the complement of $A,$ $\text{Pr}(\bar{A})=1-\text{Pr}(A)$;

4. $\text{Pr}(A \bigcup B)= \text{Pr}(A)+\text{Pr}(B)-\text{Pr}(A\bigcap B)$.

## Example: Dice

For a die the Sample Space is all the faces of the die $$S=\{1,2,3,4,5,6\}.$$ The probability of each event is $$\frac{1}{6}.$$


```{r}
Events<-c("1","2","3","4","5","6") # List of 7 numbers
## Cacluate the mean in the long form
Pr<-c(1/6,1/6,1/6,1/6,1/6,1/6)
## Cacluate the mean using the function mean
Die<-data.frame(Events,Pr)
Die
```


For the die example a Sample Space, such as $A=\{2,3\}=2 \bigcup 3$. 
Given $Pr(2 \bigcap 3)=0$, then $$\text{Pr}(2 \bigcup 3)= \text{Pr}(2)+\text{Pr}(3)=\frac{1}{6}+\frac{1}{6}.$$
