---
title: "Mathematical Probability"
author: "[John S Butler (TU Dublin)](https://johnsbutler.netlify.com)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# First Moment

## Mean (average)
### Definition:

The mean $\bar{x}$ is the point around which the sum of the deviations is 0.
The formula is given by
$$ \bar{x}=\frac{\sum_{i=1}^n x_i}{n} $$
where $x_i$ is each element and $n$ is the number of elements.
In words, add the numbers together and divided by how many numbers are in the list.

### Example
Given the list of 7 ages at a concert {19,18,20,18,18,18,20}, the mean is:

$$ \bar{x}=\frac20+18+{19+18+18+18+20}{7}=18.71429. $$


In R code:

```{r}
Age=c(20, 18,19, 18,18,18,20) # List of 7 numbers
## Cacluate the mean in the long form
(20+18+19+18+18+18+19+20)/7
## Cacluate the mean using the function mean
mean(Age) # The mean of the list

```
### Graphical representation of the mean:

```{r,warning=FALSE}
hist(Age,col="lightblue",xlab="Age (years)") # Histogram of the 7 numbers
abline(v=mean(Age),col="blue",lwd=3)# line indicating the mean
legend("topright",c("mean"),lwd=4,col=c("blue"))


```


### Mean Pros and Cons
Pros of the mean:

* The mean is easy to calculate.

* The mean is well known and understood.

Cons of the mean:
* The mean is  senstitive to extreme values.

Given the list of the ages of  8 people at a concert {20,18,19, 18,56, 18,18,20}, the mean is:

$$ \bar{x}=\frac{20+18+19+18+18+56+18+20}{8}=23.375. $$


In R code:

```{r}
Age=c(20,18,19, 18,56, 18,18,20) # List of 8 ages
## Cacluate the mean using the function mean
mean(Age) # The mean of the list

```
## Median

### Definition:

The median is the middle values of an ordered list.

### Example
Given the list of 7 ages at a concert {19,18,20,18,18,18,20}. 

1. First the list has to be ordered
  {18,18,18,18,19,20,20}.
  
2. Then the middle of the list found
  {18,18,18,*18*,19,20,20}.
  
The median is 18.


In R code:

```{r}
Age=c(20,18,19,18,18,18,20) # List of 7 numbers

median(Age)

```
### Graphical representation of the median:

```{r,warning=FALSE}
hist(Age,col="lightblue",xlab="Age (years)") # Histogram of the 7 numbers
abline(v=median(Age),col="green",lwd=3)# line indicating the mean
legend("topright",c("median"),lwd=4,col=c("green"))

```
### Median Pros and Cons
Pros of the median:

* The median is the centre of the data - very intuitive.
* Not as sensitive as the mean to extreme values.
 


Cons of median:
* The calculation of the median does not use all data.
* Sorting a long list can be computationally expensive.

## Mode

### Definition:

The mode is the most commonly occurring value in the distribution.

### Example
Given the list of 7 ages at a concert {19,18,20,18,18,18,20}. 
The most frequent values is 18, so 18 in the mode.


### In R code:

```{r}
Age=c(20,18,19,18,18,18,20) # List of 7 numbers
# Function to calculate mode
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
Mode(Age)

```
### Graphical representation of the mode:

```{r,warning=FALSE}
hist(Age,col="lightblue",xlab="Age (years)") # Histogram of the 7 numbers
abline(v=Mode(Age),col="red",lwd=3)# line indicating the mean
legend("topright",c("mode"),lwd=4,col=c("red"))

```


### Median Pros and Cons
Pros of the mode:

* The mode shows where the data is concentrated.



Cons of the mode:

* More than one Mode in a dataset.
* The calculation of the mode does not use all the data.


