---
title: "Types of Data"
author: '[John S Butler (TU Dublin)](https://johnsbutler.netlify.com)'
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

# Types of Data

1. Categorical

2. Ordinal

3. Interval

4. Ratio


## Categorical Data
### Definition:

Categorical data is the statistical data type consisting of categorical variables or of data that has been converted into that form, for example as grouped data.

### Example of Categorical Data
If you ask a number of people their age, you can categorise them into age groups. Each person can only be assigned to one age group.
This kind of data can be illustrated as a pie chart or a bar chart.


```{r, echo=FALSE}
Ages<-c("0-14yrs","15-21yrs","22-40 yrs","41-65yrs","65+ years")
N<-c(4000,3000,2000,2000,3500)
Dublin<-data.frame(N,Ages)

ggplot(Dublin, aes(x="", y=N,fill=Ages))+geom_bar(stat="identity", width=1)+ coord_polar("y", start=0)+ ggtitle("Pie Chart of Age Category")

ggplot(Dublin, aes(x=Ages, y=N,fill=Ages))+geom_bar(stat="identity")+ theme(legend.position = "none")+ ggtitle("Bar Chart of Age Category")

```

## Ordinal Data
### Definition:

Ordinal data is a categorical, statistical data type where the variables have natural, ordered categories and the distances between the categories is not known. 

## Example:
A survey asked, "Do you like Marmite?" and gave 5 possible response:

1. Strongly Disagree

2. Disagree

3. Neutral

4. Agree

5. Strongly Agree

```{r, echo=FALSE}

Marmite<-data.frame(N=c(30,10,2,10,20),Groups=factor(c("Strongly Disagree","Disagree","Neutral","Agree","Strongly Agree"),levels=c("Strongly Disagree","Disagree","Neutral","Agree","Strongly Agree")))

ggplot(Marmite, aes(x=Groups, y=N))+geom_bar(stat="identity", position=position_dodge())+ theme(legend.position = "none")+ ggtitle("Do you like Marmite?")


```

## Interval Data
### Definition:

Interval data is defined as a data type which is measured along a scale, each step in the scale has the same meaning.

## Example:
Height is a lovely example of interval data a centimeter is always a centimeter.

```{r, echo=FALSE}
Demographics<-data.frame(
                Height<-rnorm(1000,165,15),
                Weight=round(c(rnorm(200, mean=55, sd=5)))
)
ggplot(Demographics,aes(x=Height)) + geom_histogram()+ ggtitle("Height Histogram")+xlab("Height (cm)")

```

## Interval Data
### Definition:

Interval data is defined as a data type which is measured along a scale, each step in the scale has the same meaning.

## Example:
People Height Distribution

```{r, echo=FALSE}
Height<-rnorm(1000,165,15)
BMI<-rnorm(1000,20,5)
Weight<-(Height/100)*(Height/100)*BMI

Demographics<-data.frame(Height,Weight,BMI)
ggplot(Demographics,aes(x=BMI)) + geom_histogram()+ ggtitle("BMI Histogram")+xlab(bquote(Weight(kg)/(Height(m))^2))

```
